## Survival analyses

Aim: Compare with KPM analyses done by Philippe Naquet and Richard Miallot.

Todo: adapt [this tutorial](https://www.biostars.org/p/153013/)

* RFS: In cancer, the length of time after primary treatment for a cancer ends that the patient survives without any signs or symptoms of that cancer. In a clinical trial, measuring the RFS is one way to see how well a new treatment works. Also called DFS, disease-free survival, and relapse-free survival.
* OS: The length of time from either the date of diagnosis or the start of treatment for a disease, such as cancer, that patients diagnosed with the disease are still alive. In a clinical trial, measuring the OS is one way to see how well a new treatment works. Also called overall survival.



```{r resultsSurvival, cache=FALSE, echo=FALSE, warning=FALSE, results="asis"}
source("loadEnv.R")

# ggplot2 call inside survminer produces a warning.
## Warning: Vectorized input to `element_text()` is not officially supported.
## Results may be unexpected or may change in future versions of ggplot2.
knitr::opts_chunk$set(warning = FALSE)

smi <- "out/r/annotate_tcga/vsd.rds"
linkSmiToRmdAndBookDir(smi)
vsd <- readRDS(smi)

d <- colData(vsd)
rm(vsd)

d <- d[
  ,
  grepl(
    pattern = "primary_diagnosis|paper_short_histo|project_id|VNN1|status|days",
    x = colnames(d)
  )
]
d$paper_DSS_status <- as.numeric(d$paper_DSS_status)


d$days_to_death_or_last_follow_up <- ifelse(
  d$vital_status %in% "Dead",
  d$days_to_death,
  d$days_to_last_follow_up
)

d$vital_status_bin <- ifelse(
  d$vital_status %in% "Dead",
  1,
  0
)

produceDataTableWithButtons(data.frame(d))
```

### VNN1 high VS low {#resultsSurvivalAnalysisVnn1HighVsLow}

```{r resultsSurvivalAnalysisTcgaSarcPaper, fig.cap="Kaplan-Meier plot"}
# fit <- survfit(
#   Surv(
#     paper_OS_days,
#     paper_DSS_status
#   ) ~ VNN1_group_by_state,
#   data = d
# )
# ggsurvplot(
#   fit,
#   data = d,
#   pval = TRUE,
#   conf.int = TRUE,
#   risk.table = TRUE
# )
```

```{r resultsSurvivalAnalysisVnn1HighVsLowAllSelectedTcgaProjects, fig.cap="Kaplan-Meier plot for all selected TCGA projects"}
fit <- survfit(
  Surv(
    d$days_to_death_or_last_follow_up,
    d$vital_status_bin,
  ) ~ VNN1_group_by_state,
  data = d
)
ggsurvplot(
  fit,
  data = d,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

#### By project id

```{r resultsSurvivalAnalysisVnn1HighVsLowTcgaSarc, fig.cap="Kaplan-Meier plot for TCGA-SARC"}
ds <- d[d$project_id %in% "TCGA-SARC", ]
fit <- survfit(
  Surv(
    ds$days_to_death_or_last_follow_up,
    ds$vital_status_bin,
  ) ~ VNN1_group_by_project_id,
  data = ds
)
ggsurvplot(
  fit,
  data = ds,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

```{r resultsSurvivalAnalysisVnn1HighVsLowTcgaOv, fig.cap="Kaplan-Meier plot for TCGA-OV"}
ds <- d[d$project_id %in% "TCGA-OV", ]
fit <- survfit(
  Surv(
    ds$days_to_death_or_last_follow_up,
    ds$vital_status_bin,
  ) ~ VNN1_group_by_project_id,
  data = ds
)
ggsurvplot(
  fit,
  data = ds,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

```{r resultsSurvivalAnalysisVnn1HighVsLowTcgaEsca, fig.cap="Kaplan-Meier plot for TCGA-ESCA"}
ds <- d[d$project_id %in% "TCGA-ESCA", ]
fit <- survfit(
  Surv(
    ds$days_to_death_or_last_follow_up,
    ds$vital_status_bin,
  ) ~ VNN1_group_by_project_id,
  data = ds
)
ggsurvplot(
  fit,
  data = ds,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

```{r resultsSurvivalAnalysisVnn1HighVsLowTcgaPaad, fig.cap="Kaplan-Meier plot for TCGA-PAAD"}
ds <- d[d$project_id %in% "TCGA-PAAD", ]
fit <- survfit(
  Surv(
    ds$days_to_death_or_last_follow_up,
    ds$vital_status_bin,
  ) ~ VNN1_group_by_project_id,
  data = ds
)
ggsurvplot(
  fit,
  data = ds,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

```{r resultsSurvivalAnalysisVnn1HighVsLowTcgaMeso, fig.cap="Kaplan-Meier plot for TCGA-MESO"}
ds <- d[d$project_id %in% "TCGA-MESO", ]
fit <- survfit(
  Surv(
    ds$days_to_death_or_last_follow_up,
    ds$vital_status_bin,
  ) ~ VNN1_group_by_project_id,
  data = ds
)
ggsurvplot(
  fit,
  data = ds,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

```{r resultsSurvivalAnalysisVnn1HighVsLowTcgaBrca, fig.cap="Kaplan-Meier plot for TCGA-BRCA"}
ds <- d[d$project_id %in% "TCGA-BRCA", ]
fit <- survfit(
  Surv(
    ds$days_to_death_or_last_follow_up,
    ds$vital_status_bin,
  ) ~ VNN1_group_by_project_id,
  data = ds
)
ggsurvplot(
  fit,
  data = ds,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

```{r resultsSurvivalAnalysisVnn1HighVsLowTcgaLihc, fig.cap="Kaplan-Meier plot for TCGA-LIHC"}
ds <- d[d$project_id %in% "TCGA-LIHC", ]
fit <- survfit(
  Surv(
    ds$days_to_death_or_last_follow_up,
    ds$vital_status_bin,
  ) ~ VNN1_group_by_project_id,
  data = ds
)
ggsurvplot(
  fit,
  data = ds,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

```{r resultsSurvivalAnalysisVnn1HighVsLowTcgaBlca, fig.cap="Kaplan-Meier plot for TCGA-BLCA"}
ds <- d[d$project_id %in% "TCGA-BLCA", ]
fit <- survfit(
  Surv(
    ds$days_to_death_or_last_follow_up,
    ds$vital_status_bin,
  ) ~ VNN1_group_by_project_id,
  data = ds
)
ggsurvplot(
  fit,
  data = ds,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

```{r resultsSurvivalAnalysisVnn1HighVsLowTcgaStad, fig.cap="Kaplan-Meier plot for TCGA-STAD"}
ds <- d[d$project_id %in% "TCGA-STAD", ]
fit <- survfit(
  Surv(
    ds$days_to_death_or_last_follow_up,
    ds$vital_status_bin,
  ) ~ VNN1_group_by_project_id,
  data = ds
)
ggsurvplot(
  fit,
  data = ds,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

```{r resultsSurvivalAnalysisVnn1HighVsLowTcgaLusc, fig.cap="Kaplan-Meier plot for TCGA-LUSC"}
ds <- d[d$project_id %in% "TCGA-LUSC", ]
fit <- survfit(
  Surv(
    ds$days_to_death_or_last_follow_up,
    ds$vital_status_bin,
  ) ~ VNN1_group_by_project_id,
  data = ds
)
ggsurvplot(
  fit,
  data = ds,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

```{r resultsSurvivalAnalysisVnn1HighVsLowTcgaLuad, fig.cap="Kaplan-Meier plot for TCGA-LUAD"}
ds <- d[d$project_id %in% "TCGA-LUAD", ]
fit <- survfit(
  Surv(
    ds$days_to_death_or_last_follow_up,
    ds$vital_status_bin,
  ) ~ VNN1_group_by_project_id,
  data = ds
)
ggsurvplot(
  fit,
  data = ds,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

#### By primary diagnosis

```{r resultsSurvivalAnalysisVnn1HighVsLowPrimaryDiagnosisSquamousCellCarcinomaNos, fig.cap="Kaplan-Meier plot for **Squamous cell carcinoma, NOS**"}
ds <- d[d$primary_diagnosis %in% "Squamous cell carcinoma, NOS", ]
fit <- survfit(
  Surv(
    ds$days_to_death_or_last_follow_up,
    ds$vital_status_bin,
  ) ~ VNN1_group_by_primary_diagnosis,
  data = ds
)
ggsurvplot(
  fit,
  data = ds,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```
