#### Pathway {#resultsWgcnaSynthesisPathway}

The xlsx table for the qvalues of all significant Pathway gene sets is available [here](resultsWgcnaSynthesisPathway.xlsx).

```{r}
source("loadEnv.R")
```

```{r resultsWgcnaSynthesisPathway}
res_files <- list.files(
  path = file.path(
    book_from_rmd,
    "rds"
  ),
  pattern = "resultsWgcna.*Pathway.rds",
  full.names = TRUE
)
names(res_files) <- gsub(
  pattern = paste0(
    book_from_rmd,
    "/rds/resultsWgcna|KnitComputeEnrichPathway.rds"
  ),
  replacement = "",
  x = res_files
)
res_tables <- lapply(res_files, readRDS)

extractTopPvalForEachSignifTerm <- function(df) {
  d <- df@compareClusterResult[order(df@compareClusterResult$pvalue),]
  d <- d[!duplicated(d$ID),]
  d$ID <- paste(
    d$ID,
    d$Description
  )
  d <- d[, c("ID", "GeneRatio", "BgRatio", "pvalue", "qvalue")]
  return(d)
}
top_res_tables <- lapply(res_tables, extractTopPvalForEachSignifTerm)

d <- reshape2::melt(
  top_res_tables,
  id.vars = names(top_res_tables[[1]])
)
setnames(
  d,
  old = "L1",
  new = "dataset"
)
rm(res_files, res_tables, top_res_tables)

d$GeneRatio <- sapply(d$GeneRatio, function(x) eval(parse(text = x)))

# when the generatio is 1, the qvalue is NA. Better to have pvalue instead of NA for plotting
d[is.na(d$qvalue),"qvalue"] <- d[is.na(d$qvalue),"pvalue"]

fig_width <- 14

# var_facet_ont: specify a column in d to use as grouping feature for gene sets.
plotDotplotWgcnaSynthesis <- function(
  d,
  col_trans='identity',
  var_facet_ont = 'none') {
  p <- ggplot(
  d,
  aes(
    x=dataset,
    y=ID,
    color = qvalue,
    size = GeneRatio
    )
  )
  p <- p + geom_point()
  p <- p + scale_color_gradient(
    low = "red",
    high = "blue",
    trans = col_trans
  )
  p <- p + theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    )
  )
  if (var_facet_ont != "none"){
    p <- p + facet_grid(
      rows=var_facet_ont,
      scales = "free_y",
      space = "free_y",
      labeller = label_wrap_gen(width=10)
    )
  }
  p
}

n_signif_occurences <- table(d$ID)
names_signif_non_ubiquitous_occurences <- names(n_signif_occurences)[n_signif_occurences != max(n_signif_occurences)]

# Here uncommon means the hallmark is enriched in less than half the datasets
names_signif_uncommon_occurences <- names(n_signif_occurences)[n_signif_occurences < median(n_signif_occurences)]

ds <- reshape2::dcast(
  data = d,
  formula = ID~dataset,
  value.var = "qvalue"
)
wb <- createWorkbook()
sheet_name <- "Pathway"
addWorksheet(
  wb,
  sheet_name
)
writeDataTable(
  wb,
  sheet_name,
  ds,
  startCol = 1
)

freezePane(
  wb,
  sheet = sheet_name,
  firstRow = TRUE,
  firstCol = TRUE
)
headerStyle <- createStyle(
  # fontSize = 18,
  # fontName = "Arial",
  # textDecoration = "bold",
  # halign = "left",
  # fgFill = "#1A33CC",
  # border = "TopBottomLeftRight",
  textRotation = 90
)
addStyle(
  wb,
  sheet = sheet_name,
  style = headerStyle,
  rows = 1,
  cols = 1:ncol(ds)
)
setRowHeights(
  wb,
  sheet = sheet_name,
  rows = 1,
  heights = 200
)
setColWidths(
  wb,
  sheet = sheet_name,
  cols = 2:ncol(ds),
  widths = 4,
)
setColWidths(
  wb,
  sheet = sheet_name,
  cols = 1,
  widths = 32,
)
conditionalFormatting(
  wb,
  sheet = sheet_name,
  rows = 1:nrow(ds) + 1,
  type = "colourScale",
  style = c("darkgreen", "white"),
  #rule = c(1, 0),
  cols = which(names(ds) != "ID")
)
saveWorkbook(
  wb,
  file = file.path(
    book_from_rmd,
    "resultsWgcnaSynthesisPathway.xlsx"
  ),
  overwrite = TRUE
)

groups <- read.xlsx(
  file.path(
    mw_from_rmd,
    "inp/resultsWgcnaSynthesisPathwayGroupsByPN_2021-06-21.xlsx"
  )
)
d <- merge(
  d,
  groups,
  by="ID",
  all.x=TRUE
)

names_interest_Pathway <- groups[!is.na(groups$Group),"ID"]
```

##### Alphabetically-ordered Pathway and all datasets {#resultsWgcnaSynthesisPathwayAllDatasetsAlphabet}

```{r resultsWgcnaSynthesisPathwayAllDatasetsAlphabetAllSignif, dependson="resultsWgcnaSynthesisPathway", fig.width=fig_width, fig.height=120, fig.cap="Dotplot for Pathway significant (q-value < 0.05) in at least one dataset."}
plotDotplotWgcnaSynthesis(d)
```

```{r resultsWgcnaSynthesisPathwayAllDatasetsAlphabetNonUbiquitousSignif, dependson="resultsWgcnaSynthesisPathway", fig.width=fig_width, fig.height=100, fig.cap="Dotplot for Pathway which are significant (q-value < 0.05) in at least one dataset and non ubiquitous, i.e. not significant for all datasets."}
plotDotplotWgcnaSynthesis(d[d$ID %in% names_signif_non_ubiquitous_occurences,])
```

```{r resultsWgcnaSynthesisPathwayAllDatasetsAlphabetUncommonSignif, dependson="resultsWgcnaSynthesisPathway", fig.width=fig_width, fig.height=80, fig.cap="Dotplot for Pathway which are significant (q-value < 0.05) in at least one dataset, and uncommon, i.e. significant in less than half the datasets."}
plotDotplotWgcnaSynthesis(d[d$ID %in% names_signif_uncommon_occurences,])
```

```{r resultsWgcnaSynthesisPathwayAllDatasetsAlphabetInterestPathway, dependson="resultsWgcnaSynthesisPathway", fig.width=fig_width, fig.height=19, fig.cap="Dotplot for Pathway of interest (provided by Philipe Naquet)."}
plotDotplotWgcnaSynthesis(
  d[d$ID %in% names_interest_Pathway,],
  var_facet_ont = "Group"
)
```

##### Occurency-ordered Pathway and all datasets {#resultsWgcnaSynthesisPathwayAllDatasetsOccurencyOrder}

```{r resultsWgcnaSynthesisPathwayAllDatasetsOccurencyOrderAllSignif, dependson="resultsWgcnaSynthesisPathway", fig.width=fig_width, fig.height=120, fig.cap="Dotplot for Pathway significant (q-value < 0.05) in at least one dataset."}
d$ID <- factor(
  d$ID,
  levels = names(table(d$ID))[order(table(d$ID))]
)
plotDotplotWgcnaSynthesis(d)
```

```{r resultsWgcnaSynthesisPathwayAllDatasetsOccurencyOrderNonUbiquitousSignif, dependson="resultsWgcnaSynthesisPathway", fig.width=fig_width, fig.height=100, fig.cap="Dotplot for Pathway which are significant (q-value < 0.05) in at least one dataset and non ubiquitous, i.e. not significant for all datasets."}
plotDotplotWgcnaSynthesis(d[d$ID %in% names_signif_non_ubiquitous_occurences,])
```

```{r resultsWgcnaSynthesisPathwayAllDatasetsOccurencyOrderUncommonSignif, dependson="resultsWgcnaSynthesisPathway", fig.width=fig_width, fig.height=80, fig.cap="Dotplot for Pathway which are significant (q-value < 0.05) in at least one dataset, and uncommon, i.e. significant in less than half the datasets."}
plotDotplotWgcnaSynthesis(d[d$ID %in% names_signif_uncommon_occurences,])
```

```{r resultsWgcnaSynthesisPathwayAllDatasetsOccurencyOrderInterestPathway, dependson="resultsWgcnaSynthesisPathway", fig.width=fig_width, fig.height=19, fig.cap="Dotplot for Pathway of interest (provided by Philipe Naquet)."}
plotDotplotWgcnaSynthesis(
  d[d$ID %in% names_interest_Pathway,],
  var_facet_ont = "Group"
)
```

##### Log10 transformed color scale {#resultsWgcnaSynthesisPathwayAllDatasetsOccurencyOrderLog10Trans}

```{r resultsWgcnaSynthesisPathwayAllDatasetsOccurencyOrderLog10TransAllSignif, dependson="resultsWgcnaSynthesisPathway", fig.width=fig_width, fig.height=120, fig.cap="Dotplot for Pathway significant (q-value < 0.05) in at least one dataset."}
plotDotplotWgcnaSynthesis(d, col_trans = "log10")
```

```{r resultsWgcnaSynthesisPathwayAllDatasetsOccurencyOrderLog10TransNonUbiquitousSignif, dependson="resultsWgcnaSynthesisPathway", fig.width=fig_width, fig.height=100, fig.cap="Dotplot for Pathway which are significant (q-value < 0.05) in at least one dataset and non ubiquitous, i.e. not significant for all datasets."}
plotDotplotWgcnaSynthesis(
  d[d$ID %in% names_signif_non_ubiquitous_occurences,],
  col_trans = "log10"
)
```

```{r resultsWgcnaSynthesisPathwayAllDatasetsOccurencyOrderLog10TransUncommonSignif, dependson="resultsWgcnaSynthesisPathway", fig.width=fig_width, fig.height=80, fig.cap="Dotplot for Pathway which are significant (q-value < 0.05) in at least one dataset, and uncommon, i.e. significant in less than half the datasets."}
plotDotplotWgcnaSynthesis(
  d[d$ID %in% names_signif_uncommon_occurences,],
  col_trans = "log10"
)
```

```{r resultsWgcnaSynthesisPathwayAllDatasetsOccurencyOrderLog10TransInterestPathway, dependson="resultsWgcnaSynthesisPathway", fig.width=fig_width, fig.height=19, fig.cap="Dotplot for Pathway of interest (provided by Philipe Naquet)."}
plotDotplotWgcnaSynthesis(
  d[d$ID %in% names_interest_Pathway,],
  var_facet_ont = "Group",
  col_trans = "log10"
)
```

