### Neutrophil infiltration Yes Vs No {#resultsSurvivalAnalysisNeutrophilInfiltrationYesVsNo}

Currently only for TCGA-SARC (CIBERSORTx fails to compute all TCGA sample in one batch...).
Neutrophil infiltration is defined as described below:

* Yes: More than 0.4% neutrophils in sample
* No: < 0.0005% neutrophils in sample
* interm: between the two tresholds above

```{r resultsSurvivalAnalysisNeutrophilInfiltrationYesVsNo}
smi <- "inp/cibersortx/CIBERSORTx_Job3_Results.csv"
linkSmiToRmdAndBookDir(smi)
d_cibersort <- fread(smi)

ds <- d[d$project_id %in% "TCGA-SARC", ]

ds <- data.table(
  Mixture = rownames(ds),
  as.data.table(ds)
)

ds <- merge(
  ds,
  d_cibersort,
  by = "Mixture"
)

ds[
  ,
  neutrophil_infiltration := ifelse(
    ds$Neutrophils < 0.000005,
    "No",
    ifelse(
      ds$Neutrophils > 0.004,
      "Yes",
      "Int"
    )
  )
]
ds[
  ,
  neutrophil_infiltration := factor(
    neutrophil_infiltration,
    levels = c("No", "Int", "Yes")
  )
]
```

#### By project id {#resultsSurvivalAnalysisNeutrophilInfiltrationYesVsNoByProjectId}

```{r resultsSurvivalAnalysisNeutrophilInfiltrationYesVsNoByProjectId, fig.cap="Kaplan-Meier plot for TCGA-SARC"}
fit <- survfit(
  Surv(
    ds$days_to_death_or_last_follow_up,
    ds$vital_status_bin,
  ) ~ neutrophil_infiltration,
  data = ds
)
ggsurvplot(
  fit,
  data = ds,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```


#### By paper short histo {#resultsSurvivalAnalysisNeutrophilInfiltrationYesVsNoByPaperShortHisto}

##### STLMS {#resultsSurvivalAnalysisNeutrophilInfiltrationYesVsNoByPaperShortHistoSTLMS}

```{r resultsSurvivalAnalysisNeutrophilInfiltrationYesVsNoByPaperShortHistoSLMS, fig.cap="Kaplan-Meier plot for STLMS"}
dss <- ds[paper_short_histo %in% "STLMS"]
fit <- survfit(
  Surv(
    dss$days_to_death_or_last_follow_up,
    dss$vital_status_bin,
  ) ~ neutrophil_infiltration,
  data = dss
)
ggsurvplot(
  fit,
  data = dss,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

##### DDLPS {#resultsSurvivalAnalysisNeutrophilInfiltrationYesVsNoByPaperShortHistoDDLPS}

```{r resultsSurvivalAnalysisNeutrophilInfiltrationYesVsNoByPaperShortHistoDDLPS, fig.cap="Kaplan-Meier plot for DDLPS"}
dss <- ds[paper_short_histo %in% "DDLPS"]
fit <- survfit(
  Surv(
    dss$days_to_death_or_last_follow_up,
    dss$vital_status_bin,
  ) ~ neutrophil_infiltration,
  data = dss
)
ggsurvplot(
  fit,
  data = dss,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

##### UPS {#resultsSurvivalAnalysisNeutrophilInfiltrationYesVsNoByPaperShortHistoUPS}

```{r resultsSurvivalAnalysisNeutrophilInfiltrationYesVsNoByPaperShortHistoUPS, fig.cap="Kaplan-Meier plot for UPS"}
dss <- ds[paper_short_histo %in% "UPS"]
fit <- survfit(
  Surv(
    dss$days_to_death_or_last_follow_up,
    dss$vital_status_bin,
  ) ~ neutrophil_infiltration,
  data = dss
)
ggsurvplot(
  fit,
  data = dss,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```


#### VNN1 survival for non-infiltrated-by-neutrophil sarcoma

```{r resultsSurvivalAnalysisVnn1HighVsLowNonInfiltratedByNeutrophilSarcoma, fig.cap="Kaplan-Meier plot for non-infiltrated-by-neutrophil sarcoma"}
dss <- ds[neutrophil_infiltration %in% "No"]
fit <- survfit(
  Surv(
    dss$days_to_death_or_last_follow_up,
    dss$vital_status_bin,
  ) ~ VNN1_group_by_project_id,
  data = dss
)
ggsurvplot(
  fit,
  data = dss,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

```{r resultsSurvivalAnalysisVnn1TertilesNonInfiltratedByNeutrophilSarcoma, fig.cap="Kaplan-Meier plot for non-infiltrated-by-neutrophil sarcoma"}
dss <- ds[neutrophil_infiltration %in% "No"]
fit <- survfit(
  Surv(
    dss$days_to_death_or_last_follow_up,
    dss$vital_status_bin,
  ) ~ VNN1_tertiles_by_project_id,
  data = dss
)
ggsurvplot(
  fit,
  data = dss,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

#### VNN1 survival for infiltrated-by-neutrophil sarcoma
```{r resultsSurvivalAnalysisVnn1HighVsLowInfiltratedByNeutrophilSarcoma, fig.cap="Kaplan-Meier plot for infiltrated-by-neutrophil sarcoma"}
dss <- ds[neutrophil_infiltration %in% "Yes"]
fit <- survfit(
  Surv(
    dss$days_to_death_or_last_follow_up,
    dss$vital_status_bin,
  ) ~ VNN1_group_by_project_id,
  data = dss
)
ggsurvplot(
  fit,
  data = dss,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

```{r resultsSurvivalAnalysisVnn1TertilesInfiltratedByNeutrophilSarcoma, fig.cap="Kaplan-Meier plot for infiltrated-by-neutrophil sarcoma"}
dss <- ds[neutrophil_infiltration %in% "Yes"]
fit <- survfit(
  Surv(
    dss$days_to_death_or_last_follow_up,
    dss$vital_status_bin,
  ) ~ VNN1_tertiles_by_project_id,
  data = dss
)
ggsurvplot(
  fit,
  data = dss,
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE
)
```

