## Cell deconvolution with CIBSERSORTx


```{r resultsCibersortxExportValues}
# Need adjustement because now vsd also contains samples outside TCGA-SARC
# d_to_export <- data.table(
#   GeneSymbol = rowData(vsd)$hugo_symbol,
#   assay(vsd)
# )
# fwrite(
#   x = d_to_export,
#   file = file.path(
#     gdc_dir,
#     "TCGA-SARC_data_for_cibersort.tsv"
#   ),
#   sep = "\t"
# )
# rm(d_to_export)
```

The dataset is imported to CIBERSORTx and cell deconvolution is run using these parameters:

* Date: 2021-01-16 07:05:59
* Job type: Impute Cell Fractions
* Signature matrix file: LM22.update-gene-symbols.txt
* Mixture file: TCGA-SARC_data_for_cibersort.tsv
* Batch correction: enabled
* Batch correction mode: B-mode
* Source GEP file used for batch correction: LM22.update-gene-symbols.txt
* Disable quantile normalization: true
* Run mode (relative or absolute): relative
* Permutations: 100

Results correspond to Job3 from Philippe Naquet account and are retrieved and available in the [mw-sarcoma-data repository](https://github.com/guillaumecharbonnier/mw-sarcoma-data/tree/main/inp/cibersortx) (Private until publication).

```{r resultsCibersortxImportResults}
source("loadEnv.R")
smi <- "out/r/annotate_tcga/vsd.rds"
linkSmiToRmdAndBookDir(smi)
vsd <- readRDS(smi)

smi <- "inp/cibersortx/CIBERSORTx_Job3_Results.csv"
linkSmiToRmdAndBookDir(smi)
d_cibersort <- fread(smi)

d <- data.table(
  VNN1_VST_signal = assay(vsd)[
    rowData(vsd)$hugo_symbol == "VNN1",
    d_cibersort$Mixture
  ],
  d_cibersort
)

d_metadata_samples <- as.data.table(
  colData(vsd)[
    ,
    c(
      "primary_diagnosis",
      "paper_short_histo",
      "barcode"
    )
  ]
)

d <- merge(
  d,
  d_metadata_samples,
  by.x = "Mixture",
  by.y = "barcode"
)

d$Mixture <- factor(
  d$Mixture,
  levels = d$Mixture[order(d$VNN1_VST_signal)]
)
```


```{r resultsCibersortxPlotVnn1, fig.height=35, fig.width=8, fig.cap="This barplot is a visual explanation for the ordering of samples used for other barplots below."}
p <- ggplot(
  d,
  aes(
    fill = VNN1_VST_signal,
    x = Mixture,
    y = VNN1_VST_signal
  )
)
p <- p + geom_col()
#  stat = "identity"
# )
p <- p + coord_flip()
p <- p + theme(
  legend.position = "top",
  legend.justification = "left"
)
p
```


```{r resultsCibersortxPlotStack, fig.height=35, fig.width=8, fig.cap="Stacked barplots for CIBERSORTx results. Fancy, but not really readable. See barplots below."}
d <- melt(
  d,
  id.vars = c(
    "VNN1_VST_signal",
    "Mixture",
    "P-value",
    "Correlation",
    "RMSE",
    "primary_diagnosis",
    "paper_short_histo"
  ),
  variable.name = "Cell_type",
  value.name = "Proportion"
)
p <- ggplot(
  d,
  aes(
    fill = Cell_type,
    y = Proportion,
    x = Mixture
  )
)
p <- p + geom_bar(
  position = "stack",
  stat = "identity"
)
p <- p + coord_flip()
p <- p + theme(
  legend.position = "top",
  legend.justification = "left"
)
p
```


```{r resultsCibersortxPlotFacet, fig.height=35, fig.width=16, fig.cap="Same CIBERSORTx results but splitted by cell types on facets."}
p <- ggplot(
  d,
  aes(
    fill = Cell_type,
    x = Mixture,
    y = Proportion
  )
)
p <- p + facet_grid(~Cell_type, scale = "free")
p <- p + geom_bar(
  # position = "stack",
  stat = "identity"
)
p <- p + coord_flip()
p <- p + theme(
  legend.position = "top"
)
p
```

```{r resultsCibersortxPlotFacetOnlyCellTypesCorrelatedWithVnn1, fig.height=35, fig.width=12, fig.cap="Focus on cell types that should express VNN1, or correlate with it."}
selected_cell_types <- c(
  "Monocytes",
  "Macrophages M0",
  "Macrophages M1",
  "Macrophages M2",
  "Neutrophils",
  "T cells CD8"
)
p <- ggplot(
  d[Cell_type %in% selected_cell_types],
  aes(
    fill = Cell_type,
    x = Mixture,
    y = Proportion
  )
)
p <- p + facet_grid(~Cell_type, scale = "free")
p <- p + geom_bar(
  # position = "stack",
  stat = "identity"
)
p <- p + coord_flip()
p <- p + theme(
  legend.position = "top"
)
p
```

```{r resultsCibersortxPlotFacetOnlyCellTypesCorrelatedWithVnn1ByPrimaryDiagnosis, fig.height=45, fig.width=8, fig.cap="Focus on cell types that should express VNN1. Facet by primary diagnosis. For each facet, samples are still ordered from high-to-low VNN1 expression from top-to-bottom."}
p <- p + facet_grid(
  primary_diagnosis ~ Cell_type,
  scales = "free",
  space = "free_y"
)
p
```

```{r resultsCibersortxPlotFacetOnlyCellTypesCorrelatedWithVnn1ByPaperShortHisto, fig.height=45, fig.width=8, fig.cap="Focus on cell types that should express VNN1. Facet by paper short histo."}
p <- p + facet_grid(
  paper_short_histo ~ Cell_type,
  scales = "free",
  space = "free_y"
)
p
```



```{r resultsCibersortxPlotFacetOnlyCellTypesCorrelatedWithVnn1Stack, fig.height=35, fig.width=8, fig.cap="Focus on cell types that should express VNN1. Stacked view allow  to see the cell type distribution match VNN1 signal overall, but with a few strong exceptions, *eg* TCGA-HB-A2OT-01A-11R-A21T-07."}
p <- ggplot(
  d[Cell_type %in% selected_cell_types],
  aes(
    fill = Cell_type,
    x = Mixture,
    y = Proportion
  )
)
p <- p + geom_bar(
  position = "stack",
  stat = "identity"
)
p <- p + coord_flip()
p <- p + theme(
  legend.position = "top"
)
p
```

```{r resultsCibersortxPlotFacetOnlyCellTypesCorrelatedWithVnn1StackByPrimaryDiagnosis, fig.height=45, fig.width=8, fig.cap="Focus on cell types that should express VNN1. Stacked view. Facet by primary diagnosis."}
p <- p + facet_grid(
  primary_diagnosis ~ .,
  scales = "free",
  space = "free"
)
p
```

```{r resultsCibersortxPlotFacetOnlyCellTypesCorrelatedWithVnn1StackByPaperShortHisto, fig.height=45, fig.width=8, fig.cap="Focus on cell types that should express VNN1. Stacked view. Facet by paper short histo."}
p <- p + facet_grid(
  paper_short_histo ~ .,
  scales = "free",
  space = "free"
)
p
```

### Highlight Neutrophils

```{r resultsCibersortxCorrelationNeutroVNN1, fig.height=7, fig.width=8, fig.cap="Correlation between neutrophil proportion and VNN1 signal."}
d_neutro <- d[Cell_type %in% "Neutrophils"]
lm_eqn <- function(df) {
  m <- lm(Proportion ~ VNN1_VST_signal, df)
  eq <- substitute(
    italic(y) == a + b %.% italic(x) * "," ~ ~ italic(r)^2 ~ "=" ~ r2,
    list(
      a = format(coef(m)[1], digits = 2),
      b = format(coef(m)[2], digits = 2),
      r2 = format(summary(m)$r.squared, digits = 3)
    )
  )
  as.character(as.expression(eq))
}
eq <- lm_eqn(d_neutro)
p <- ggplot(
  data = d_neutro,
  aes(
    x = VNN1_VST_signal,
    y = Proportion,
    color = paper_short_histo
  )
)
p <- p + geom_smooth(
  method = "lm",
  se = FALSE,
  color = "black",
  formula = y ~ x
)
p <- p + geom_point()
p <- p + geom_text(
  aes(
    x = 6,
    y = 0.01,
    label = eq
  ),
  parse = TRUE,
  inherit.aes = FALSE
)
p
```


```{r resultsCibersortxCorrelationNeutroVNN1BySarcomaSubtypes, fig.height=17, fig.width=8, fig.cap="Correlation between neutrophil proportion and VNN1 signal, by sarcoma subtypes"}
eq <- ddply(
  d_neutro,
  .(paper_short_histo),
  lm_eqn
)
p <- ggplot(
  data = d_neutro,
  aes(
    x = VNN1_VST_signal,
    y = Proportion
  )
)
p <- p + geom_smooth(
  method = "lm",
  se = FALSE,
  color = "black",
  formula = y ~ x
)
p <- p + geom_point(color = "grey")
p <- p + geom_text(
  data = eq,
  aes(
    x = 6,
    y = 0.01,
    label = V1
  ),
  parse = TRUE,
  inherit.aes = FALSE
)
p <- p + facet_grid(paper_short_histo ~ .)
p
```
