#### Kegg {#resultsWgcnaSynthesisKegg}

The xlsx table for the qvalues of all significant Kegg gene sets is available [here](resultsWgcnaSynthesisKegg.xlsx).

```{r}
source("loadEnv.R")
```

```{r resultsWgcnaSynthesisKegg}
res_files <- list.files(
  path = file.path(
    book_from_rmd,
    "rds"
  ),
  pattern = "resultsWgcna.*Kegg.rds",
  full.names = TRUE
)
names(res_files) <- gsub(
  pattern = paste0(
    book_from_rmd,
    "/rds/resultsWgcna|KnitComputeEnrichKegg.rds"
  ),
  replacement = "",
  x = res_files
)
res_tables <- lapply(res_files, readRDS)

extractTopPvalForEachSignifTerm <- function(df) {
  d <- df@compareClusterResult[order(df@compareClusterResult$pvalue),]
  d <- d[!duplicated(d$ID),]
  d$ID <- paste(
    d$ID,
    d$Description
  )
  d <- d[, c("ID", "GeneRatio", "BgRatio", "pvalue", "qvalue")]
  return(d)
}
top_res_tables <- lapply(res_tables, extractTopPvalForEachSignifTerm)

d <- reshape2::melt(
  top_res_tables,
  id.vars = names(top_res_tables[[1]])
)
setnames(
  d,
  old = "L1",
  new = "dataset"
)
rm(res_files, res_tables, top_res_tables)

d$GeneRatio <- sapply(d$GeneRatio, function(x) eval(parse(text = x)))

# when the generatio is 1, the qvalue is NA. Better to have pvalue instead of NA for plotting
d[is.na(d$qvalue),"qvalue"] <- d[is.na(d$qvalue),"pvalue"]

fig_width <- 9

plotDotplotWgcnaSynthesis <- function(
  d,
  col_trans='identity',
  var_facet_ont = 'none',
  var_facet_dataset = 'none',
  dataset_label = "dataset"
) {
  d <- droplevels(d)
  p <- ggplot(
  d,
  aes_string(
    x = dataset_label,
    y = "ID",
    color = "qvalue",
    size = "GeneRatio"
    )
  )
  p <- p + geom_point()
  p <- p + scale_color_gradient(
    low = "red",
    high = "blue",
    trans = col_trans
  )
  p <- p + theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      vjust = 1
    )
  )
  if (var_facet_ont != "none"){
    var_facet_ont <- sym(var_facet_ont)
    p <- p + facet_grid(
      rows = vars(!!var_facet_ont),
      scales = "free_y",
      space = "free_y",
      labeller = label_wrap_gen(width=10)
    )
  }
  if (var_facet_dataset != "none" & var_facet_ont != "none"){
    # https://stackoverflow.com/questions/67983042/why-passing-a-string-in-r-ggplot2facet-grid-works-for-rows-but-not-for-colum
    var_facet_dataset <- sym(var_facet_dataset)
    var_facet_ont <- sym(var_facet_ont)
    p <- p + facet_grid(
      rows = vars(!!var_facet_ont),
      cols = vars(!!var_facet_dataset),
      scales = "free",
      space = "free",
      labeller = label_wrap_gen(width=30)
    )
  }
  p
}

n_signif_occurences <- table(d$ID)
names_signif_non_ubiquitous_occurences <- names(n_signif_occurences)[n_signif_occurences != max(n_signif_occurences)]

# Here uncommon means the hallmark is enriched in less than half the datasets
names_signif_uncommon_occurences <- names(n_signif_occurences)[n_signif_occurences < median(n_signif_occurences)]

ds <- reshape2::dcast(
  data = d,
  formula = ID~dataset,
  value.var = "qvalue"
)
wb <- createWorkbook()
sheet_name <- "Kegg"
addWorksheet(
  wb,
  sheet_name
)
writeDataTable(
  wb,
  sheet_name,
  ds,
  startCol = 1
)

freezePane(
  wb,
  sheet = sheet_name,
  firstRow = TRUE,
  firstCol = TRUE
)
headerStyle <- createStyle(
  # fontSize = 18,
  # fontName = "Arial",
  # textDecoration = "bold",
  # halign = "left",
  # fgFill = "#1A33CC",
  # border = "TopBottomLeftRight",
  textRotation = 90
)
addStyle(
  wb,
  sheet = sheet_name,
  style = headerStyle,
  rows = 1,
  cols = 1:ncol(ds)
)
setRowHeights(
  wb,
  sheet = sheet_name,
  rows = 1,
  heights = 200
)
setColWidths(
  wb,
  sheet = sheet_name,
  cols = 2:ncol(ds),
  widths = 4,
)
setColWidths(
  wb,
  sheet = sheet_name,
  cols = 1,
  widths = 32,
)
conditionalFormatting(
  wb,
  sheet = sheet_name,
  rows = 1:nrow(ds) + 1,
  type = "colourScale",
  style = c("darkgreen", "white"),
  #rule = c(1, 0),
  cols = which(names(ds) != "ID")
)
saveWorkbook(
  wb,
  file = file.path(
    book_from_rmd,
    "resultsWgcnaSynthesisKegg.xlsx"
  ),
  overwrite = TRUE
)

groups <- read.xlsx(
  file.path(
    mw_from_rmd,
    "inp/KeggPhilippeNaquetGroups.xlsx"
  )
)
d <- merge(
  d,
  groups,
  by="ID",
  all.x=TRUE
)

# Use manual order from Philippe Naquet
d$ID <- factor(
  d$ID,
  levels = rev(groups$ID)
)
d$Group <- factor(
  d$Group,
  levels = unique(groups$Group)
)
datasets_groups <- read.xlsx(
  file.path(
    mw_from_rmd,
    "inp/dataset_id_to_label.xlsx"
  )
)
d <- merge(
  d,
  datasets_groups,
  by="dataset",
  all.x=TRUE
)

# Use manual order from Philippe Naquet
d$label <- factor(
  d$label,
  levels = datasets_groups$label
)
d$dataset_type <- factor(
  d$dataset_type,
  levels = unique(datasets_groups$dataset_type)
)

names_interest_Kegg <- groups[!is.na(groups$Group),"ID"]
names_interest_datasets <- datasets_groups[!is.na(datasets_groups$label),"label"]
```

##### Manually-selected and ordered pathways and datasets {#resultsWgcnaSynthesisKeggAllDatasetsManual}

Below are the two variants with either linear or log scale for q-values.

```{r resultsWgcnaSynthesisKeggAllDatasetsAlphabetInterestKegg, dependson="resultsWgcnaSynthesisKegg", fig.width=fig_width, fig.height=13, fig.cap="Dotplot for Kegg of interest (provided by Philipe Naquet)."}
plotDotplotWgcnaSynthesis(
  d[d$ID %in% names_interest_Kegg & d$label %in% names_interest_datasets,],
  var_facet_ont = "Group",
  var_facet_dataset = "dataset_type",
  dataset_label = "label"
)
```

```{r resultsWgcnaSynthesisKeggAllDatasetsOccurencyOrderLog10TransInterestKegg, dependson="resultsWgcnaSynthesisKegg", fig.width=fig_width, fig.height=13, fig.cap="Dotplot for Kegg of interest (provided by Philipe Naquet)."}
plotDotplotWgcnaSynthesis(
  d[d$ID %in% names_interest_Kegg & d$label %in% names_interest_datasets,],
  var_facet_ont = "Group",
  var_facet_dataset = "dataset_type",
  dataset_label = "label",
  col_trans = "log10"
)
```

##### Alphabetically-ordered Kegg and all datasets {#resultsWgcnaSynthesisKeggAllDatasetsAlphabet}

Adding facets and manual reordering in dotplots above ruined the dotplots below. Since the first ones should be enough for the review, no effort is planned to fix the dotplots below.



```{r resultsWgcnaSynthesisKeggAllDatasetsAlphabetAllSignif, dependson="resultsWgcnaSynthesisKegg", fig.width=fig_width, fig.height=38, fig.cap="Dotplot for Kegg significant (q-value < 0.05) in at least one dataset."}
plotDotplotWgcnaSynthesis(d)
```

```{r resultsWgcnaSynthesisKeggAllDatasetsAlphabetNonUbiquitousSignif, dependson="resultsWgcnaSynthesisKegg", fig.width=fig_width, fig.height=28, fig.cap="Dotplot for Kegg which are significant (q-value < 0.05) in at least one dataset and non ubiquitous, i.e. not significant for all datasets."}
plotDotplotWgcnaSynthesis(d[d$ID %in% names_signif_non_ubiquitous_occurences,])
```

```{r resultsWgcnaSynthesisKeggAllDatasetsAlphabetUncommonSignif, dependson="resultsWgcnaSynthesisKegg", fig.width=fig_width, fig.height=19, fig.cap="Dotplot for Kegg which are significant (q-value < 0.05) in at least one dataset, and uncommon, i.e. significant in less than half the datasets."}
plotDotplotWgcnaSynthesis(d[d$ID %in% names_signif_uncommon_occurences,])
```


##### Occurency-ordered Kegg and all datasets {#resultsWgcnaSynthesisKeggAllDatasetsOccurencyOrder}

```{r resultsWgcnaSynthesisKeggAllDatasetsOccurencyOrderAllSignif, dependson="resultsWgcnaSynthesisKegg", fig.width=fig_width, fig.height=38, fig.cap="Dotplot for Kegg significant (q-value < 0.05) in at least one dataset."}
d$ID <- factor(
  d$ID,
  levels = names(table(d$ID))[order(table(d$ID))]
)
plotDotplotWgcnaSynthesis(d)
```

```{r resultsWgcnaSynthesisKeggAllDatasetsOccurencyOrderNonUbiquitousSignif, dependson="resultsWgcnaSynthesisKegg", fig.width=fig_width, fig.height=28, fig.cap="Dotplot for Kegg which are significant (q-value < 0.05) in at least one dataset and non ubiquitous, i.e. not significant for all datasets."}
plotDotplotWgcnaSynthesis(d[d$ID %in% names_signif_non_ubiquitous_occurences,])
```

```{r resultsWgcnaSynthesisKeggAllDatasetsOccurencyOrderUncommonSignif, dependson="resultsWgcnaSynthesisKegg", fig.width=fig_width, fig.height=19, fig.cap="Dotplot for Kegg which are significant (q-value < 0.05) in at least one dataset, and uncommon, i.e. significant in less than half the datasets."}
plotDotplotWgcnaSynthesis(d[d$ID %in% names_signif_uncommon_occurences,])
```

```{r resultsWgcnaSynthesisKeggAllDatasetsOccurencyOrderInterestKegg, dependson="resultsWgcnaSynthesisKegg", fig.width=fig_width, fig.height=5, fig.cap="Dotplot for Kegg of interest (provided by Philipe Naquet)."}
plotDotplotWgcnaSynthesis(
  d[d$ID %in% names_interest_Kegg & d$label %in% names_interest_datasets,],
  var_facet_ont = "Group",
  var_facet_dataset = "dataset_type"
)
```

##### Log10 transformed color scale {#resultsWgcnaSynthesisKeggAllDatasetsOccurencyOrderLog10Trans}

```{r resultsWgcnaSynthesisKeggAllDatasetsOccurencyOrderLog10TransAllSignif, dependson="resultsWgcnaSynthesisKegg", fig.width=fig_width, fig.height=38, fig.cap="Dotplot for Kegg significant (q-value < 0.05) in at least one dataset."}
plotDotplotWgcnaSynthesis(d, col_trans = "log10")
```

```{r resultsWgcnaSynthesisKeggAllDatasetsOccurencyOrderLog10TransNonUbiquitousSignif, dependson="resultsWgcnaSynthesisKegg", fig.width=fig_width, fig.height=28, fig.cap="Dotplot for Kegg which are significant (q-value < 0.05) in at least one dataset and non ubiquitous, i.e. not significant for all datasets."}
plotDotplotWgcnaSynthesis(
  d[d$ID %in% names_signif_non_ubiquitous_occurences,],
  col_trans = "log10"
)
```

```{r resultsWgcnaSynthesisKeggAllDatasetsOccurencyOrderLog10TransUncommonSignif, dependson="resultsWgcnaSynthesisKegg", fig.width=fig_width, fig.height=19, fig.cap="Dotplot for Kegg which are significant (q-value < 0.05) in at least one dataset, and uncommon, i.e. significant in less than half the datasets."}
plotDotplotWgcnaSynthesis(
  d[d$ID %in% names_signif_uncommon_occurences,],
  col_trans = "log10"
)
```

